Let’s Disco – Publishing person-level data as Linked Data.

### 2015-01-08 paper/ESWC 2015 Tutorial/ESWC2015Tutorial.tex
material and bio

%
%
\title{Let’s Disco – Publishing person-level data as Linked Data.}
\subtitle{Building a Data Portal with DCAT, DDI-RDF Discovery, PHDD, and other vocabularies.\bz{Ben's Kommentar}
}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXX\inst{1} \and XXX\inst{1} \and Joachim Wackerow\inst{1}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{thomas.bosch,benjamin.zapilko,joachim.wackerow\}}@gesis.org}

@@

\keywords{DDI-RDF Discovery Vocabulary, Disco, Microdata, Microdata Metadata, Discovery of Microdata, Data Documentation Initiative, DDI, Linked Data, Semantic Web}
\end{abstract}
%

@@

The DDI (Data Documentation Initiative) is a structured metadata standard related to the observation and measurement of human activity.
It started out in the mid-1990s as a replacement for traditional archival code books documenting research data, and then branched off to cover the research data life cycle. Over time, as the data landscape has
changed, the DDI XML specifications have evolved to add new coverage and functionality to respond to new user requirements.
While the Data Documentation Initiative (DDI) is an international metadata standard with origins in the quantitative social sciences, it is increasingly being used by researchers and practitioners in other disciplines. The DDI specifications are also being used to document other data types, such as social media, biomarkers, administrative data, and transaction data. The specification itself is modular and can document and manage different stages of the data lifecycle, such as conceptualization, collection, processing, analysis, distribution, discovery, repurposing, and archiving.
Existing DDI-XML instances can be transformed into this RDF format and therefore exposed as Linked Data. This workshop aims to present the Disco vocabulary and its practical appliance with DDI metadata in detail.

@@

@@

\subsection{Learning Objectives}
The learning objectives of this tutorial are the following:
\begin{itemize} 
\item Introduction of Disco for modelling complex person-level micro data
\item Representing the interplay of Disco data with existing vocabularies
\item Validation of Disco
\end{itemize}

@@

Thomas Bosch and Benjamin Zapilko held a half-day tutorial\footnote{tutorial slides avaliable at: \url{http://de.slideshare.net/boschthomas/201412-lets-disco-eddi-2014} and \url{http://de.slideshare.net/boschthomas/201412-lets-disco-2-eddi-2014}} at the EDDI14 – 6th Annual European DDI User Conference in London\footnote{\url{http://www.eddi-conferences.eu/ocs/index.php/eddi/eddi14/schedConf/program}}.
The tutorial was well received and the presenters got overall positive feedback. Since the audience of disco is two-fold - data professionals and the DDI community at the one hand and Linked Data experts at the other hand - the idea was to offer this tutorial at the ESWC conference as well.
At the EDDI conference, the first part of the audience - data professionals and the DDI community - was addressed.
At the ESWC conference, the second part of our audience - Linked Data experts - will be addressed.

@@

>>>>

%
%
\title{Let’s Disco – Publishing person-level data as Linked Data.}
\subtitle{Building a Highly-Complex Data Set with Disco, PHDD, XKOS and widely used vocabularies.
}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXX\inst{1} \and XXX\inst{1} \and Joachim Wackerow\inst{1}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{thomas.bosch,benjamin.zapilko,joachim.wackerow\}}@gesis.org}

@@

\keywords{DDI-RDF Discovery Vocabulary, Disco, Microdata Metadata, Data Documentation Initiative, Linked Data}
\end{abstract}
%

@@

The DDI (Data Documentation Initiative) is a structured metadata standard related to the observation and measurement of human activity.
It started out in the mid-1990s as a replacement for traditional archival code books documenting research data, and then branched off to cover the research data life cycle. Over time, as the data landscape has
changed, the DDI XML specifications have evolved to add new coverage and functionality to respond to new user requirements.
While the Data Documentation Initiative (DDI) is an international metadata standard with origins in the quantitative social sciences, it is increasingly being used by researchers and practitioners in other disciplines. The DDI specifications are also being used to document other data types, such as social media, biomarkers, administrative data, and transaction data. The specification itself is modular and can document and manage different stages of the data lifecycle, such as conceptualization, collection, processing, analysis, distribution, discovery, repurposing, and archiving.
This workshop aims to present the Disco vocabulary and its practical appliance with DDI metadata in detail.

@@

During the tutorial, we built one real world complex data set which serves as ongoing running example. 
The whole data set can be exported from the Microdata Information System\footnote{\url{http://www.gesis.org/missy/editor/}}. 
In all sessions, there will be practical hands-on exercises where the participants are asked to model and query particular parts of this data set. 
The infrastructure for these exercises, i.e. a triple store, will be provided for all participants.
We will extend slides of a previous tutorial\footnote{tutorial slides avaliable at: \url{http://de.slideshare.net/boschthomas/201412-lets-disco-eddi-2014} and \url{http://de.slideshare.net/boschthomas/201412-lets-disco-2-eddi-2014}} 
and where appropriate refer to the current Disco specification\footnote{current specification is available at \url{https://github.com/linked-statistics/disco-spec}}.

@@

\subsection{Learning Objectives}
The learning objectives of this tutorial are the following:
\begin{itemize} 
\item Introduction of Disco, PHDD, and XKOS for modelling complex person-level micro data
\item Representing the interplay of Disco data with existing vocabularies
\item Validation of Disco
\end{itemize}

@@

Thomas Bosch and Benjamin Zapilko held a half-day tutorial at the EDDI14 – 6th Annual European DDI User Conference in London\footnote{\url{http://www.eddi-conferences.eu/ocs/index.php/eddi/eddi14/schedConf/program}}.
The tutorial was well received and the presenters got overall positive feedback. Since the audience of disco is two-fold - data professionals and the DDI community at the one hand and Linked Data experts at the other hand - the idea was to offer this tutorial at the ESWC conference as well.
At the EDDI conference, the first part of the audience - data professionals and the DDI community - was addressed.
At the ESWC conference, the second part of our audience - Linked Data experts - will be addressed.

@@

Joachim Wackerow holds a degree in sociology. 
He is the vice chair of the DDI Alliance Technical Committee whose charge is to create, maintain, and enhance the DDI specification.
He is the chair of the DDI Alliance RDF Vocabularies Working Group whose charge is to develop RDF vocabularies for efficient use of DDI metadata in the context of the Semantic Web/Linked Data.
He organizes the annual European DDI user conferences and holds DDI introduction workshops.
